Labor Market Income and Substitution Effects Explain Contemporary Natural Selection

### 2020-12-20 why-natural-selection.Rmd
Add, but don't use, inequality calculation for parents' generation

Although the direction of selection effects does not change between the
generations, there are other differences. Compared to our standard polygenic
scores (residualized on 100 principal components of genetic data) 
selection effects on unresidualized scores are about ten percent higher
on average for *number of siblings*, whereas effects for *number of children*
barely change (Appendix Figure \@ref(fig:plot-regs-controlled-pcs)). This could
be because earlier fertility is driven more by geographically clustered
deprivation [e.g. via an insurance motive, @rendall1998old], which may correlate
with the broad-scale genetic variation captured by principal components, such as
ancestry.

@@

Because selection effects are concentrated in poorer groups, they may also increase
inequality with respect to polygenic scores. For example, Figure
\@ref(fig:plot-mean-EA3-by-income) graphs mean polygenic scores for educational
attainment (EA3), for children in households of different income groups. The
blue bars show the actual means, i.e. parents' mean polygenic score weighted by
number of children. The grey bars show the hypothetical means if all households
had equal numbers of children. Natural selection against genes associated with 
educational attainment, which lowers the mean, is stronger at the bottom of the 
income distribution, and this increases the differences between groups.  Overall,
natural selection increases inequality for 29 out of 33 polygenic scores 
(Appendix Table \@ref(tab:tbl-inequality-effects)).

>>>>

Although the direction of selection effects does not change between the
generations, there are other differences. Compared to our standard polygenic
scores (residualized on 100 principal components of genetic data), 
selection effects on unresidualized scores are about ten percent higher
on average for *number of siblings*, whereas effects for *number of children*
barely change (Appendix Figure \@ref(fig:plot-regs-controlled-pcs)). This could
be because earlier fertility is driven more by geographically clustered
deprivation [e.g. via an insurance motive, @rendall1998old], which may correlate
with the broad-scale genetic variation captured by principal components, such as
ancestry.

@@

Because selection effects are concentrated in poorer groups, they may also increase
inequality with respect to polygenic scores. For example, Figure
\@ref(fig:plot-mean-EA3-by-income) graphs mean polygenic scores for educational
attainment (EA3) among children from households of different income groups. The
blue bars show the actual means, i.e. parents' mean polygenic score weighted by
number of children. The grey bars show the hypothetical means if all households
had equal numbers of children. Natural selection against genes associated with 
educational attainment, which lowers the mean, is stronger at the bottom of the 
income distribution, and this increases the differences between groups.  Overall,
natural selection increases inequality for 29 out of 33 polygenic scores 
(Appendix Table \@ref(tab:tbl-inequality-effects)).
